Relationships in the extremes and their influence on competition and coexistence

### 2022-07-12 paper/Paper.Rnw
stopping edits for now

The other mechanisms in MCT that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, 
(see eqn. (13) in Ellner (2019),
whereas ATAs are %defined in a covariation context
strictly about covariation. 
Therefore, the storage effect is the only mechanism of coexistence
for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
coexistence operating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- the storage effect. 
Only storage effect involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is the storage effect. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of MCT which was goal G1 of the Introduction. One typically (but see below) must examine this quantity for each species, $i$,
to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.

>>>>

The other mechanisms in MCT that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, 
(see eqn. (13) in Ellner (2019),
whereas ATAs are %defined in a covariation context
strictly about covariation. 
Therefore, the storage effect is the only mechanism of coexistence
for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
which equals GWR of species $i$. The decomposition, which we use below, is analogous to and motivated by the 
original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
coexistence operating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- the storage effect. 
Only storage effect involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is the storage effect. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of MCT which was goal G1 of the Introduction. One typically %(but see below) 
must examine this quantity for each species, $i$,
to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.
